---
title: "SDS 322E Project"
author: "Team 2"
date: "2022-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is our project for team 2, Flight Experience. The data contains information over 129,880 people who flew on different airplanes (with the same airline company though) and describes their plane ride experience. They rate their overall experience (from a scale of 1 to 5 where '0' stands for "Not Applicable" , '1' is "Least Satisfied" and '5' is "Most Satisfied") over a lot of different variables. There are also two columns that include data over their departure delay and arrival delay time. It's a classification problem, and we will look at the variables that make the best ride for a passenger and determine if they would have categorized their ride as 'satisfied' or ' 'unsatisfied' depending on their self-report. 

Our goal: With this information, we will be able to help the company increase their retention rate with customers by being able to focus on what aspects of the flights that people enjoy the most. 

References/Link to the data: The data is given by an airline organization. The actual name of the company is not given due to privacy. This is a link to the original dataset on Kaggle: https://www.kaggle.com/datasets/sjleshrac/airlines-customer-satisfaction

These are all the column names: 

 [1] "satisfaction"                      "Gender"                            "Customer.Type"                     "Age"                              
 [5] "Type.of.Travel"                    "Class"                             "Flight.Distance"                   "Seat.comfort"                     
 [9] "Departure.Arrival.time.convenient" "Food.and.drink"                    "Gate.location"                     "Inflight.wifi.service"            
[13] "Inflight.entertainment"            "Online.support, "                    "Ease.of.Online.booking"            "On.board.service"                 
[17] "Leg.room.service"                  "Baggage.handling"                  "Checkin.service"                   "Cleanliness"                      
[21] "Online.boarding"                   "Departure.Delay.in.Minutes"        "Arrival.Delay.in.Minutes"  




```{r}
library(ggplot2)
library(tidyverse)
library(cluster)
library(vtable)
```


##Reading in the excel file
```{r}
flight_df <- read.csv("Flight_Experience.csv") %>% data.frame()
head(flight_df)
```

##Getting column information
```{r}
flight_df %>% glimpse()
```


##Determining how many nulls there are
```{r}

nulls <- flight_df %>% summarize_all(function(x) sum(is.na(x)))
t(nulls)
```
There are NA's in the dataset so lets clean it. Only column that has blanks is Arrival.Delay.in.Minutes


##Let's create a summary statstics table
```{r}
num <- flight_df %>% select(where(is.numeric)) 
st(num)
```


##Let's determine the overall satisfaction of the passengers
```{r}

flight_df %>% ggplot(aes(x = satisfaction, fill = satisfaction)) + geom_bar(position="stack", stat="count") + geom_text(stat="count", aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.2) + ggtitle("Overall Satisfaction Count")

```
##Variables that have the highest rating for zero
```{r}
df <- flight_df[ , 8:21]

df_rating <- df %>% summarise_all(function(x){sum(x==0)}) %>% 
  gather(metric, zerocount) %>% filter(zerocount > 100)


ggplot(df_rating,aes(x=metric,y=zerocount)) + geom_bar(stat="identity") + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Rating Variable Counts with 0")
```



```{r, fig.height=8, fig.width= 8}
df <- flight_df %>% select_if(is.numeric) %>% na.omit()
cormat <- df %>% cor()


cormat %>% as.data.frame %>% rownames_to_column("var1")


tidycor <- cormat %>%
  as.data.frame %>%
  rownames_to_column("var1") %>%
  pivot_longer(-1, names_to = "var2", values_to = "correlation")

  
tidycor %>% ggplot(aes(var1, var2, fill=correlation)) + 
  geom_tile() + 
  scale_fill_gradient2(low="red", mid="white", high = "blue") +
  geom_text(aes(label=round(correlation,2)),color = "black", size = 2)+ #overlays correlation values
  theme(axis.text.x = element_text(angle = 90)) + #flips the x-axis labels
  coord_fixed() + ggtitle("Correlation Between all Rating Variables")
```


Correlation map between satisfaction and rating variables
```{r,fig.height=8, fig.width= 8}

flight_df_numeric <- flight_df %>% mutate(satisfy = ifelse(satisfaction == "satisfied", 1, 0))





df2 <- flight_df_numeric %>% select_if(is.numeric) %>% na.omit()
cormat2 <- df2 %>% cor()


cormat2 %>% as.data.frame %>% rownames_to_column("var1")


tidycor2 <- cormat2 %>%
  as.data.frame %>%
  rownames_to_column("var1") %>%
  pivot_longer(-1, names_to = "var2", values_to = "correlation")
#Get every column except column 1
  
tidycor2 <- tidycor2 %>% filter(var2 == "satisfy")

tidycor2 %>% ggplot(aes(var1, var2, fill=correlation)) + 
  geom_tile() + 
  scale_fill_gradient2(low="red", mid="white", high = "blue") +
  geom_text(aes(label=round(correlation,2)),color = "black", size = 2)+ #overlays correlation values
  theme(axis.text.x = element_text(angle = 90)) + #flips the x-axis labels
  coord_fixed() + ggtitle("Correlation Between Satisfaction and Rating Variables")
```



Creating correlation map of all the demographic/independent variables
```{r, fig.height=5, fig.width= 5}



flight_df_independent <- flight_df %>% mutate(satisfy = ifelse(satisfaction == "satisfied", 1, 0), gender = ifelse(Gender == "Male", 0, 1), customer.type = ifelse(Customer.Type == "Loyal Customer", 1, 0), travel.type = ifelse(Type.of.Travel == "Personal Travel", 0, 1), class = ifelse(Class == "Eco", 0, ifelse(Class == "Business", 2, 1)))

flight_df_independent <- flight_df_independent %>% select(satisfy, gender, customer.type, travel.type, class)

df3 <- flight_df_independent %>% select_if(is.numeric) %>% na.omit()
cormat3 <- df3 %>% cor()


cormat3 %>% as.data.frame %>% rownames_to_column("var1")


tidycor3 <- cormat3 %>%
  as.data.frame %>%
  rownames_to_column("var1") %>%
  pivot_longer(-1, names_to = "var2", values_to = "correlation")
#Get every column except column 1
  

tidycor3 %>% ggplot(aes(var1, var2, fill=correlation)) + 
  geom_tile() + 
  scale_fill_gradient2(low="red", mid="white", high = "blue") +
  geom_text(aes(label=round(correlation,2)),color = "black", size = 4)+ #overlays correlation values
  theme(axis.text.x = element_text(angle = 90)) + #flips the x-axis labels
  coord_fixed() + ggtitle("Correlation Between Demographic Variables")

#male 0, female 1
#business 2, eco plus 1, eco 0
#personal 0, business travel 1
#loyal 1, disloyal 0
#satisfy 1, disatisfy 0


```



Looking at all the numeric variables and their average rating to satisfaction
```{r}
m <- function(x){
  
  mean(x, rm.na = T)
  
}
average <- flight_df %>% group_by(satisfaction) %>% select_if(is.numeric) %>% summarize_all(m) %>% select(-Flight.Distance)
#average


tidy_average <- average %>% pivot_longer(cols=c('Seat.comfort':'Online.boarding'), names_to = 'flight_variable', values_to = 'average')
#tidy_average

tidy_average %>%  ggplot(aes(x = satisfaction, y = average, fill = satisfaction)) + geom_bar(stat = "identity") + facet_wrap(~flight_variable) + ggtitle("Rating Variables by Satisfaction")

difference_df <- tidy_average %>% select(flight_variable, satisfaction, average) %>% arrange(flight_variable, average) %>% mutate(diff=average-lag(average,default=first(average))) %>% mutate(difference = ifelse(diff <= 0, "-", diff)) %>% select(-diff)


difference_df %>% filter(difference != "-") %>% ggplot(aes(x = flight_variable, y=difference, color = satisfaction)) + geom_point()  + geom_segment(aes(x=flight_variable, xend=flight_variable, y=0, yend=difference)) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Difference in Satisfaction Average")

```



By gender

```{r}

ggplot(data = flight_df, aes(x = Gender, fill = satisfaction)) + geom_bar(position="dodge") +  geom_text(stat="count", aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.2) + theme_bw() + ggtitle("Satisfaction by Gender")

```

By customer type
```{r}
ggplot(data = flight_df, aes(x = Customer.Type, fill = satisfaction)) + geom_bar(position="dodge") +  geom_text(stat="count", aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.2) + theme_bw() + ggtitle("Satisfaction by Customer Type")

```

By type of travel
```{r}

ggplot(data = flight_df, aes(x = Type.of.Travel , fill = satisfaction)) + geom_bar(position="dodge") +  geom_text(stat="count", aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.2) + theme_bw() + ggtitle("Satisfaction by Type of Travel")
```

By class
```{r}
ggplot(data = flight_df, aes(x = Class , fill = satisfaction)) + geom_bar(position="dodge") +  geom_text(stat="count", aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.2) + theme_bw() + ggtitle("Satisfaction by Class")
```



Let's create some categorical variables for age
```{r}


flight_df %>% ggplot(aes(x = Age, y= satisfaction, fill = satisfaction)) + geom_violin() + ggtitle("Age and Satisfaction")
flight_df %>% ggplot(aes(x = Age, y= satisfaction, fill = satisfaction)) + geom_violin() + facet_wrap(~Gender) + ggtitle("Age and Sex")


```


Clustering
```{r}

flight_df %>% ggplot(aes(x = Age, y = Departure.Delay.in.Minutes, color = satisfaction)) + geom_point(size = .2) + facet_wrap(~satisfaction) + ggtitle("Age and Departure Delay")
#flight_df %>% ggplot(aes(x = Age, y = Flight.Distance, color = satisfaction)) + geom_point(size = .2) + facet_wrap(~satisfaction)

clust_dat <- flight_df %>% select(Age, Departure.Delay.in.Minutes)

kmeans1 <- clust_dat %>% kmeans(4) #set number of clusters k

kmeansclust <- clust_dat %>% mutate(cluster=as.factor(kmeans1$cluster)) 



```



```{r}

kmeansclust %>% ggplot(aes(x = Age, y = Departure.Delay.in.Minutes, color = cluster)) + geom_point(size = .2) + ggtitle("Clustering Between Age and Departure Delay")
```

Doing some data analysis to to determine the clustering and the class they belong to
```{r}
join_cluster <- flight_df %>% select(satisfaction, Age, Departure.Delay.in.Minutes, Class) 

temp <- inner_join(join_cluster, kmeansclust, by = c("Age", "Departure.Delay.in.Minutes")) %>% distinct() %>% group_by(cluster, Class) %>% summarize(satisfy = sum(satisfaction == "satisfied"), nosatisfy = sum(satisfaction == "dissatisfied")) %>% arrange(desc(nosatisfy))
temp

```



Analyzing our clustering soultion
```{r}

wss<-vector() #create empty vector

#for k=1, 2, ... 10, compute WSS and save it in wss
for(i in 1:10){
  temp2 <- flight_df %>% select(Age,Departure.Delay.in.Minutes) %>% kmeans(i)
  wss[i] <- temp2$tot.withinss # tot.withinss = WSS
}

wss
ggplot()+geom_point(aes(x=1:10,y=wss))+geom_path(aes(x=1:10,y=wss))+
  xlab("clusters")+scale_x_continuous(breaks=1:10) + ggtitle("WSS to find Optimal Clustern Amount")
```



More visuzlations about class type and flight distance
```{r}



ggplot(flight_df, aes(x = Class, y = Flight.Distance, fill = satisfaction)) + geom_bar(stat = "summary", fun=mean, position="dodge") + geom_errorbar(stat="summary", fun.data = mean_se) + facet_wrap(~satisfaction) + theme_bw() + ggtitle("Class Type and Flight Distance")




```


Leg room and type of travel
```{r}


flight_df %>% ggplot(aes(x = Leg.room.service, fill = Type.of.Travel)) + geom_bar(stat = "count") + facet_wrap(~satisfaction) + ggtitle("Leg Room by Type of Travel")



```


Gender overall distribution 
```{r}
flight_df %>% group_by(Gender) %>% summarize(male = sum(Gender == "Male"), female = sum(Gender == "Female"))
data <- data.frame(
  category=c("Male", "Female"),
  count=c(63981, 65899)
)

# Compute percentages
data$fraction <- data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax <- cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin <- c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(data$category, "\n value: ", data$count)

# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=6) +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) + theme_void() +
  theme(legend.position = "none") + ggtitle("General Gender Distribution")

```


Leg room count by gender
```{r}

flight_df %>%group_by(`Leg.room.service`,Gender)%>%summarise(count=n())
ggplot(flight_df, aes(x=Gender,y=`Leg.room.service`,fill=`Leg.room.service`))+geom_bar(stat="identity")

ggplot(flight_df,aes(x=Gender,y=Leg.room.service,fill=satisfaction))+
  geom_bar(stat="identity")+ facet_wrap(~Leg.room.service)  + ggtitle("Leg Room by Gender Count")




flights_gender <- flight_df %>% mutate(gender = ifelse(Gender == "Male", 0, 1))



df3 <- flights_gender %>% select_if(is.numeric) %>% na.omit()
cormat3 <- df3 %>% cor()



cormat3 %>% as.data.frame %>% rownames_to_column("var1")


tidycor3 <- cormat3 %>%
  as.data.frame %>%
  rownames_to_column("var1") %>%
  pivot_longer(-1, names_to = "var2", values_to = "correlation")
#Get every column except column 1
  
tidycor3 <- tidycor3 %>% filter(var2 == "gender")



tidycor3 %>% ggplot(aes(var1, var2, fill=correlation)) + 
  geom_tile() + 
  scale_fill_gradient2(low="red", mid="white", high = "blue") +
  geom_text(aes(label=round(correlation,2)),color = "black", size = 2)+ #overlays correlation values
  theme(axis.text.x = element_text(angle = 90)) + #flips the x-axis labels
  coord_fixed() + ggtitle("Correlation Between Gender and Rating Variables")

```

Correlation map for male flyers
```{r}


flight_df_gender <- flight_df %>% mutate(Gender = ifelse(Gender == "Male", 0, 1)) %>% mutate(satisfaction = ifelse(satisfaction == "satisfied", 1, 0)) %>% filter(Gender == 0)
head(flight_df_gender)


df2 <- flight_df_gender %>% select_if(is.numeric) %>% na.omit()
cormat2 <- df2 %>% cor()



cormat2 %>% as.data.frame %>% rownames_to_column("var1")



tidycor2 <- cormat2 %>%
  as.data.frame %>%
  rownames_to_column("var1") %>%
  pivot_longer(-1, names_to = "var2", values_to = "correlation")
#Get every column except column 1
  
tidycor2 <- tidycor2 %>% filter(var2 == "satisfaction") %>% filter (var1=="Inflight.entertainment" | var1=="Leg.room.service"
            | var1=="Online.boarding"
            | var1=="Online.support"
            | var1=="Seat.comfort"
            | var1=="On.board.service"
            | var1=="Inflight.wifi.service")

tidycor2 %>% ggplot(aes(var1, var2, fill=correlation)) + 
  geom_tile() + 
  scale_fill_gradient2(low="red", mid="white", high = "blue") +
  geom_text(aes(label=round(correlation,2)),color = "black", size = 4)+ #overlays correlation values
  theme(axis.text.x = element_text(angle = 90)) + #flips the x-axis labels
  coord_fixed() + ggtitle("Correlation Between Male Flyers")


```

Correlation map for female flyers

```{r}


flight_df_gender <- flight_df %>% mutate(Gender = ifelse(Gender == "Male", 0, 1)) %>% mutate(satisfaction = ifelse(satisfaction == "satisfied", 1, 0)) %>% filter(Gender == 1)
head(flight_df_gender)


df2 <- flight_df_gender %>% select_if(is.numeric) %>% na.omit()
cormat2 <- df2 %>% cor()



cormat2 %>% as.data.frame %>% rownames_to_column("var1")


tidycor2 <- cormat2 %>%
  as.data.frame %>%
  rownames_to_column("var1") %>%
  pivot_longer(-1, names_to = "var2", values_to = "correlation")
#Get every column except column 1
  
tidycor2 <- tidycor2 %>% filter(var2 == "satisfaction") %>% filter (var1=="Inflight.entertainment" | var1=="Leg.room.service"
            | var1=="Online.boarding"
            | var1=="Online.support"
            | var1=="Seat.comfort"
            | var1=="On.board.service"
            | var1=="Inflight.wifi.service")

tidycor2 %>% ggplot(aes(var1, var2, fill=correlation)) + 
  geom_tile() + 
  scale_fill_gradient2(low="red", mid="white", high = "blue") +
  geom_text(aes(label=round(correlation,2)),color = "black", size = 4)+ #overlays correlation values
  theme(axis.text.x = element_text(angle = 90)) + #flips the x-axis labels
  coord_fixed() + ggtitle("Correlation Between Female Flyers")

```



Categorizing Age
```{r}
flight_df$Age_group <- cut(flight_df$Age, 
                        breaks = c(6, 12, 19, 39, 59, 85), 
                        labels = c("Children","Teenager","Young Adult","Middle Age Adult",
                                   "Senior Adult"))
flight_df %>% ggplot(aes(x=Age_group)) + geom_bar(stat="count") + ggtitle("Age groupings Distribution")
```


Categorizing Age by intervals of 10
```{r}
flight_df$Age_group2 <- cut(flight_df$Age, 
                        breaks = c(0, 9, 19, 29, 39, 49, 59, 69, 79, 89), 
                        labels = c("below 10", "10s","20s","30s","40s", "50s", "60s", "70s", "80s"))

flight_df %>% ggplot(aes(x=Age_group2)) + geom_bar(stat="count") + ggtitle("Age groupings by interval of 10s")
```


```{r}
flight_df %>% 
  ggplot(aes(x = Age_group, fill=satisfaction)) + 
  geom_bar(position = "fill") + ylab("Satisfaction") + ggtitle("Satisfaction and Age Grouping")
```


Categorize departure and arrival delay variables
```{r, fig.height=8, fig.width= 8}
flight_df$DepartureDelay_group <- cut(flight_df$Departure.Delay.in.Minutes, 
                        breaks = c(-1, 0, 10, 20, 30, 60, 120, 180, 1600), 
                        labels = c("Punctual", "1~10min late", "11~20min late", "21~30min late", 
                                   "31~60min late", "61~120min late", "121~180min late", 
                                   "more than 180min late"))

flight_df$ArrivalDelay_group <- cut(flight_df$Arrival.Delay.in.Minutes, 
                        breaks = c(-1, 0, 10, 20, 30, 60, 120, 180, 1600), 
                        labels = c("Punctual", "1~10min late", "11~20min late", "21~30min late", 
                                   "31~60min late", "61~120min late", "121~180min late", 
                                   "more than 180min late"))

# change satisfaction variable to numeric variable(0,1)
flight2 <- flight_df %>% mutate(satisfaction=ifelse(satisfaction=="satisfied",1,0))

# Make a dataframe that shows satisfaction ratio grouped by age group and delay group
flight_ratio <- flight2 %>% 
  group_by(Age_group, DepartureDelay_group) %>% 
  summarize(sat_ratio = sum(satisfaction)/n()) %>%
  mutate(sat_ratio = round(sat_ratio, digit=2))

flight_ratio %>% 
  ggplot(aes(x=DepartureDelay_group, y=sat_ratio)) + 
  geom_bar(stat='identity', fill = "skyblue") +
  facet_wrap(~Age_group, ncol=5) + 
  geom_text(aes(label = sat_ratio), vjust = -0.3, size=3.3) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Departure Delay Time") + ylab("Satisfaction Ratio") +
  ylim(0,1) + ggtitle("Departure Delay Time and Satisfacation Ratio")
```


```{r, fig.height=8, fig.width= 8}
# change satisfaction variable to numeric variable(0,1)
flight2 <- flight_df %>% mutate(sat=ifelse(satisfaction=="satisfied",1,0))

# Make a dataframe that shows satisfaction ratio grouped by age group and delay group
flight_ratio2 <- flight2 %>% 
  group_by(Age_group, Ease.of.Online.booking) %>% 
  summarize(sat_ratio = sum(sat)/n()) %>%
  mutate(sat_ratio = round(sat_ratio, digit=2))
## `summarise()` has grouped output by 'Age_group'. You can override using the
## `.groups` argument.
flight_ratio2 %>% 
  ggplot(aes(x=Ease.of.Online.booking, y=sat_ratio)) + 
  geom_bar(stat='identity', fill = "skyblue") +
  facet_wrap(~Age_group, ncol=5) + 
  geom_text(aes(label = sat_ratio), vjust = -0.3, size=3.3) +
  theme_minimal() + 
  #theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  ylab("Satisfaction Ratio") +
  scale_x_continuous(name = "Ease in Online Booking", breaks = seq(0, 5, by = 1)) +
  ylim(0,1) + ggtitle("Ease of Online Booking and Satisfaction Ratio")

```




